---
title: "SURVMETH 745 HW10"
author: "Stacey Frank & Chendi Zhao"
date: "4/16/2022"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(PracTools)
library(survey)
library(tidyverse)
library(janitor)
library(sampling)
library(survey)
```

# 14.1. Use the smho.N874 dataset to complete this exercise on poststratification.
Select a simple random sample of size n = 80 without replacement.
If you use R, set the random number seed to −530049348 with the set.seed
command.
```{r}
set.seed(-530049348)
dat<-smho.N874
dim(dat)
names(dat)

sam<-dat[sample(nrow(dat), size=80,replace = F), ]

```
## (a) 
```{r}
dat%>%group_by(hosp.type)%>%
  summarise_at(vars(EXPTOTAL), list(mean = mean))
```
We want to check whether the means of expenditures vary across the five hospital types. If so, that means hospital type is a good predictor of the expenditures as well as a good auxiliary variable for poststratification. 
According to the output above, the means are different and we can use hospital type for poststratification.


## (b)
```{r}
tabyl(dat$hosp.type,exclude = NULL)
tabyl(sam$hosp.type,exclude = NULL)
```
From the outputs above, all five hospital types are presented in the sample. If there is one type missing, that will result in unstable estimates of the population controls and adds unnecessarily
to the variability of the final weights. To avoid these problems, we may consider collasping two types and form less strata.

## (c) 
```{r}
d <- rep(nrow(dat)/80, 80)
f1 <- rep(80/nrow(dat), 80)
srs.dsgn <- svydesign(ids = ~0, strata = NULL, fpc = ~f1, data = sam,weights = ~d)

N.PS <- xtabs(~hosp.type,data = dat)
ps.dsgn <- postStratify(design = srs.dsgn,
strata = ~hosp.type,
population = N.PS)

# Weights-Before
sum(weights(srs.dsgn))
# Weights-After
sum(weights(ps.dsgn))
```
The weights sum to before and after poststratification are the same. As we expected, the both weights sum should equal to the population
counts.

## (d)
```{r}
svytotal(~ as.factor(hosp.type), ps.dsgn, na.rm=TRUE)
```
## (e) 
```{r}
 options(scipen = n)
svytotal(~EXPTOTAL, srs.dsgn)
svytotal(~EXPTOTAL, ps.dsgn)
```
In this sample, the srs and poststratified estimated totals
are similar and the latter has slightly a smaller SE.

# 14.3 GREG weights

```{r}
data(smho.N874)
smho <- smho.N874[smho.N874$hosp.type !=4,]

# Draw section 14.3.2 sample

x <- smho[,"BEDS"]
x[x <= 5] <- 5
x <- sqrt(x)
n <- 80
set.seed(428274453)
pk <- n*x/sum(x)
sam <- UPrandomsystematic(pk)
sam <- sam==1
sam.dat <- smho[sam, ]
d <- 1/pk[sam]

smho.dsgn <- svydesign(ids = ~0, strata = NULL, data = data.frame(sam.dat), weights = ~d)

# aux pop totals

x.beds <- sum(smho$BEDS)
x.seen <- sum(smho$SEENCNT)
x.eoy <- sum(smho$EOYCNT)
# htype.count <- smho %>% count(hosp.type)
# x.htype <- htype.count$n
x.htype <- table(smho$hosp.type)
N <- nrow(smho)

pop.tots <- c('(Intercept)' = N, 
              BEDS = x.beds, 
              SEENCNT = x.seen, 
              EOYCNT = x.eoy, 
              HTYPE = x.htype)

# GREG

sam.lin <- calibrate(design = smho.dsgn,
                     formula = ~ BEDS + SEENCNT + EOYCNT + as.factor(hosp.type), 
                     population = pop.tots,
                     calfn="linear")

```

## (a) Verify that the weights are calibrated

```{r}

svytotal(~BEDS, sam.lin)
svytotal(~SEENCNT, sam.lin)
svytotal(~EOYCNT, sam.lin)

svytable(~hosp.type, sam.lin)
```

Looking for standard errors that are close to zero to demonstrate that weights are calibrated. 

## (b) What are the ranges of the base weights and calibrated weights?

```{r}

summary(weights(smho.dsgn))

summary(weights(sam.lin))

```

