---
title: "HW2_cd"
author: "Group Sarndal: Stacey Frank & Chendi Zhao"
date: "1/27/2022"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(PracTools)
library(tidyverse)
library(janitor)
library(knitr)
```

## 3.7

a)
```{r}
nPropMoe(moe.sw=1, e=.005, alpha=0.05, pU=0.02)

```
The sample size required to estimate a rare characteristic with a prevalence of 2% with a MOE of .005 will be 3012. 


b) You can assess the probability of collecting no cases with the rare trait for a variety of same sizes and prevalence estimates by solving the following inequality: $n\geq\frac{log(\alpha)}{log(1-p_u)}$. This would allow you to determine the prevalence level you could reasonably expect to detect at a specified confidence level at various sample sizes. 

## 3.8

b)
```{r, results='hide'}
data("smho98")
dim(smho98)
head(smho98)
```


```{r}
#The unit relvariance is the population relvariance, which is the square of the population coefficient of variation; formula on pg 26 

attach(smho98)
EXPTOTAL.RV <- var(EXPTOTAL)/(mean(EXPTOTAL)^2)
BEDS.RV <- var(BEDS)/(mean(BEDS)^2)
SEENCNT.RV <- var(SEENCNT)/(mean(SEENCNT)^2)
EOYCNT.RV <- var(EOYCNT)/(mean(EOYCNT)^2)
Y_IP.RV <- var(Y_IP)/(mean(Y_IP)^2)

Relvars <- cbind(EXPTOTAL.RV, BEDS.RV, SEENCNT.RV, EOYCNT.RV, Y_IP.RV)
print(Relvars)

```
The unit relvariances for each of the five variables are listed above.

## 3.12
```{r}
#See example 3.16 in VDK P66-67
#x<-c(70,160,227,235,275,300,310,378,400,461,472,474,498,562,584,635,670,712,760,816)
#y<-c(244,402,732,925,632,557,1226,896,2190,1948,1239,1258,1657,2116,1326,1606,1707,2089,1283,1239)
#sample<-cbind.data.frame(x,y)
#n=20

data(hospital)
sample<-read.table("Hospital_50.txt",header = T,sep = ",")
sample<-rename(sample,x=BEDS,y=DISCHARGE)

N=393
n=50
N_bed=107956
n_bed=sum(sample$x)
#a
p_i<-sample$x/N_bed
w_i=1/(p_i*n)
#verify
sum(w_i*sample$x)
round(w_i,digit=2)
```
#a
If the design weights are correct, the sum of number of beds in each sampled hosptital*weight should be the total number of beds in the population(107956)
```{r}
#b
y_bar_bed=round(sum(sample$y/(p_i))/(N*n),2)
y_bar_bed
```
The mean of discharges based on the sample using the $\pi$-estimator is  742.43
.

```{r}
#c
V_hat_1_c=n^2*sum((w_i*sample$y-mean(w_i*sample$y))^2)/(n-1)
V_hat_1_c=sum((sample$y/p_i-mean(sample$y/p_i))^2)/(n-1)

V_y_bar_wr=V_hat_1_c/(N^2*n)
V_y_bar_wr
```
The estimated variance of the mean of discharges is 593.062

Observation:???? The variance is small???

```{r}
#e
V_hat_1_e=sum(sqrt(hospital$x))*sum(((sample$y)^2/(sqrt(sample$x)*p_i)))/n-(sum(sample$y/p_i)/n)^2+V_y_bar_wr*N^2
V_hat_1_e
N_new=V_hat_1_e/((N*mean(hospital$y)*0.15)^2)
ceiling(N_new)
```
If we select a new sample with probabilities proportional to the square of root of beds. The appropriate $V_1$ in this case would be 37938877108 And to meet the target CV, we only need to sample 17 hospitals.


Q1
ceiling(nPropMoe(moe.sw=1,e=0.005,alpha=0.05,pU=0.02))
Q2
data("smho98")
y1<-smho98$EXPTOTAL
y_S1<-var(smho98$EXPTOTAL)
y_bar1<-mean(smho98$EXPTOTAL)
cv1<-y_S1/(y_bar1)^2
cv1

