---
title: "SURVMETH 745 HW5"
author: "Stacey Frank & Chendi Zhao"
date: "2/3/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(PracTools)
library(tidyverse)
```

# Exercise 9.1

```{r}

data("MDarea.pop")
dim(MDarea.pop)
table(MDarea.pop$TRACT)
summary(MDarea.pop)
head(MDarea.pop)

```

```{r}
###Tract as PSU

MDarea.pop %>%
  group_by(TRACT) %>%
  summarise(N = n(), y = sum(y1)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Tract Popluation") + ylab("Tract Total: y1")

MDarea.pop %>%
  group_by(TRACT) %>%
  summarise(N = n(), y = sum(y2)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Tract Popluation") + ylab("Tract Total: y2")

MDarea.pop %>%
  group_by(TRACT) %>%
  summarise(N = n(), y = sum(y3)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Tract Popluation") + ylab("Tract Total: y3")

MDarea.pop %>%
  group_by(TRACT) %>%
  summarise(N = n(), y = sum(ins.cov)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Tract Popluation") + ylab("Tract Total: ins.cov")

MDarea.pop %>%
  group_by(TRACT) %>%
  summarise(N = n(), y = sum(hosp.stay)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Tract Popluation") + ylab("Tract Total: hosp.stay")

```


```{r}

###Block group as PSU

MDarea.pop %>%
  group_by(10*TRACT + BLKGROUP) %>%
  summarise(N = n(), y = sum(y1)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Block Group Popluation") + ylab("Block Group Total: y1")

MDarea.pop %>%
  group_by(10*TRACT + BLKGROUP) %>%
  summarise(N = n(), y = sum(y2)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Block Group Popluation") + ylab("Block Group Total: y2")

MDarea.pop %>%
  group_by(10*TRACT + BLKGROUP) %>%
  summarise(N = n(), y = sum(y3)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Block Group Popluation") + ylab("Block Group Total: y3")

MDarea.pop %>%
  group_by(10*TRACT + BLKGROUP) %>%
  summarise(N = n(), y = sum(ins.cov)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Block Group Popluation") + ylab("Block Group Total: ins.cov")

MDarea.pop %>%
  group_by(10*TRACT + BLKGROUP) %>%
  summarise(N = n(), y = sum(hosp.stay)) %>%
  ggplot(aes(x=N, y=y)) +
  geom_point(size=2) +
  xlab("Block Group Popluation") + ylab("Block Group Total: hosp.stay")

```

These plots show that there is a strong linear relationship between the population count for the PSU (be it tract or block group) and the cluster total for each of the five variables in the Maryland dataset. As the population size increases, the cluster total for each of the five variables increases as well. This creates large between-cluster variation and makes the B^2 value very large. A good way to counteract this is to have clusters that are roughly equal in size, but absent that solution, it is better to do a PPS selection of clusters, rather than an SRS. This gives larger clusters a larger probability of selection and thereby decreases the between-cluster variation that will be found in the resulting sample.

You can see the decrease in B^2 values below when comparing an SRS selection of tracts and a PPS selection of tracts. The B^2 value is lower in the PPS setup for all five variables. 


#If TRACT is defined as PSU
##Assume an srs/srs sampling design
```{r}

MDarea.pop%>%
  group_by(TRACT)%>%
  summarise(N=n())%>%
  summary(N)

TRACT1<-round(rbind(y1=BW2stageSRS(MDarea.pop$y1, psuID=MDarea.pop$TRACT),
            y2=BW2stageSRS(MDarea.pop$y2, psuID=MDarea.pop$TRACT),
            y3=BW2stageSRS(MDarea.pop$y3, psuID=MDarea.pop$TRACT),
            ins.cov=BW2stageSRS(MDarea.pop$ins.cov, psuID=MDarea.pop$TRACT),
            hosp.stay=BW2stageSRS(MDarea.pop$hosp.stay, psuID=MDarea.pop$TRACT)),
      5)
```

##Assume a ppswr/srs sampling design

```{r}
pp.TRACT <- table(MDarea.pop$TRACT) / nrow(MDarea.pop)
TRACT2<-round(rbind(y1=BW2stagePPS(MDarea.pop$y1, pp=pp.TRACT, psuID=MDarea.pop$TRACT),
            y2=BW2stagePPS(MDarea.pop$y2, pp=pp.TRACT, psuID=MDarea.pop$TRACT),
            y3=BW2stagePPS(MDarea.pop$y3, pp=pp.TRACT, psuID=MDarea.pop$TRACT),
            ins.cov=BW2stagePPS(MDarea.pop$ins.cov, pp=pp.TRACT, psuID=MDarea.pop$TRACT),
            hosp.stay=BW2stagePPS(MDarea.pop$hosp.stay, pp=pp.TRACT, psuID=MDarea.pop$TRACT)),
      5)
```

```{r}
kable(list(TRACT1,TRACT2),caption = "TRACT as PSU--srs/srs&ppswr/srs")
```
#If TRACT is defined as PSU
##Assume an srs/srs sampling design
```{r}
MDarea.pop$BP<-10*MDarea.pop$TRACT + MDarea.pop$BLKGROUP
MDarea.pop%>%
  group_by(BP)%>%
  summarise(N=n())%>%
  summary(N)

BP1<-round(rbind(y1=BW2stageSRS(MDarea.pop$y1, psuID=MDarea.pop$BP),
            y2=BW2stageSRS(MDarea.pop$y2, psuID=MDarea.pop$BP),
            y3=BW2stageSRS(MDarea.pop$y3, psuID=MDarea.pop$BP),
            ins.cov=BW2stageSRS(MDarea.pop$ins.cov, psuID=MDarea.pop$BP),
            hosp.stay=BW2stageSRS(MDarea.pop$hosp.stay, psuID=MDarea.pop$BP)),
      5)
```

##Assume a ppswr/srs sampling design

```{r}
pp.BP <- table(MDarea.pop$BP) / nrow(MDarea.pop)
BP2<-round(rbind(y1=BW2stagePPS(MDarea.pop$y1, pp=pp.BP, psuID=MDarea.pop$BP),
            y2=BW2stagePPS(MDarea.pop$y2, pp=pp.BP, psuID=MDarea.pop$BP),
            y3=BW2stagePPS(MDarea.pop$y3, pp=pp.BP, psuID=MDarea.pop$BP),
            ins.cov=BW2stagePPS(MDarea.pop$ins.cov, pp=pp.BP, psuID=MDarea.pop$BP),
            hosp.stay=BW2stagePPS(MDarea.pop$hosp.stay, pp=pp.BP, psuID=MDarea.pop$BP)),
      5)
```

```{r}
kable(list(BP1,BP2),caption = "BLOCKGROUP as PSU--srs/srs&ppswr/srs")
```

If we set TRACT or BLOCKGROUP as PSU,the $B^2$ for all five variables are smaller under a ppswr design than an srs design.The approximation of $S^2/U^2$ (unit variance) = $B^2+W^2$ works better under a ppswr design. That's also can be concluded from that $k$ is close to 1 for all variables under a ppswr design, indicating a homogeneity within each cluster.
Therefore, we think that pps sampling would be more efficient
