---
title: "SURVMETH 745 HW6"
author: "Stacey Frank & Chendi Zhao"
date: "2/26/2022"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(PracTools)
library(tidyverse)
```

#Exercise 9.7
##Data
```{r}
data("MDarea.pop")
table(MDarea.pop$Hispanic,exclude = NULL)
table(MDarea.pop$Gender,exclude = NULL)
summary(MDarea.pop$Age)
```

## a (three-stage(all srswr))
```{r}
M <- length(unique(MDarea.pop$TRACT))
pp.trt1 <- rep(1/M,M)
MDarea.pop$trtBG <- 10*MDarea.pop$TRACT + MDarea.pop$BLKGROUP
res1<-round(rbind(
  y1=BW3stagePPS(X=MDarea.pop$Hispanic, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y2=BW3stagePPS(X=MDarea.pop$Gender, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y3=BW3stagePPS(X=MDarea.pop$Age, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG)),digits = 3)
rownames(res1)<-c("Hispanic","Gender","Age")
res1

```

## b (three-stage(ppswr/srs/srs))
```{r}

pp.trt2 <- table(MDarea.pop$TRACT) / nrow(MDarea.pop) 
MDarea.pop$trtBG <- 10*MDarea.pop$TRACT + MDarea.pop$BLKGROUP
res2<-round(rbind(
  y1=BW3stagePPS(X=MDarea.pop$Hispanic, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y2=BW3stagePPS(X=MDarea.pop$Gender, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y3=BW3stagePPS(X=MDarea.pop$Age, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG)),digits = 3)

rownames(res2)<-c("Hispanic","Gender","Age")
res2
```


##c
For all three variables,when swithcing to a ppswr/srs/srs design to an srs/srs/srs design.$W^2$,$w2^2$,and $W3^2$ decreased slightly and $\delta2^2$ increased slgihtly. However, the value of $B^2$ and $\delta1^2$ became much smaller than before. The $\delta1^2$ is a measure of the homogeneity among the PSU totals. Decrease in this term indicates that the variation
within PSUs is much larger than the variation among PSU totals under the ppswr/srs/srs design. This is most obvious in the varible "Hispanis","Gender".
