---
title: "SURVMETH 745 HW6"
author: "Stacey Frank & Chendi Zhao"
date: "2/26/2022"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(PracTools)
library(tidyverse)
```
# Exercise 9.5
```{r}
#total budget: 500,000
#cost: c(1000, 200, 120)
#delta1: c(.001, .01, .05)
#delta2: c(.05, .10, .25)
#pi estimator is used so we're doing srs/srs/srs

#delta1=.001
Optimum1<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .001, delta2= .05, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum2<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .001, delta2= .10, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum3<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .001, delta2= .25, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)

#delta1=.01
Optimum4<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .01, delta2= .05, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum5<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .01, delta2= .10, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum6<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .01, delta2= .25, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)

#delta1=.05
Optimum7<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .05, delta2= .05, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum8<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .05, delta2= .10, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)
Optimum9<- clusOpt3(unit.cost=c(1000, 200, 120), delta1= .05, delta2= .25, unit.rv=2, k1=1, k2=1, tot.cost=500000, cal.sw=1)

delta1<- c(.001,.001,.001, .01,.01,.01, .05, .05, .05)
delta2<- c(.05, .10, .25,.05, .10, .25,.05, .10, .25)
m.opt <- c(Optimum1[[10]], Optimum2[[10]], Optimum3[[10]],Optimum4[[10]], Optimum5[[10]], Optimum6[[10]],Optimum7[[10]], Optimum8[[10]], Optimum9[[10]])
n.opt <- c(Optimum1[[11]], Optimum2[[11]], Optimum3[[11]],Optimum4[[11]], Optimum5[[11]], Optimum6[[11]],Optimum7[[11]], Optimum8[[11]], Optimum9[[11]])
q.opt <- c(Optimum1[[12]], Optimum2[[12]], Optimum3[[12]],Optimum4[[12]], Optimum5[[12]], Optimum6[[12]],Optimum7[[12]], Optimum8[[12]], Optimum9[[12]])
CV <-  c(Optimum1[[13]], Optimum2[[13]], Optimum3[[13]],Optimum4[[13]], Optimum5[[13]], Optimum6[[13]],Optimum7[[13]], Optimum8[[13]], Optimum9[[13]])

design.options <- data.frame(delta1, delta2, m.opt, n.opt, q.opt, CV)

design.options
```
The m.opt, n.opt, q.opt, and CV calues for each of the nine combinations of delta1 and delta2 are displayed above. CV is smallest when delta1 and delta2 values are small. Smaller delta1 values are associated with smaller m.opt values, and smaller delta2 values are associated with smaller n.opt values. q.opt gets smaller as delta2 increases (within a given delta1 value).

# Exercise 9.7
## Data
```{r}
data("MDarea.pop")
table(MDarea.pop$Hispanic,exclude = NULL)
table(MDarea.pop$Gender,exclude = NULL)
summary(MDarea.pop$Age)
```

## a (three-stage(all srswr))
```{r}
M <- length(unique(MDarea.pop$TRACT))
pp.trt1 <- rep(1/M,M)
MDarea.pop$trtBG <- 10*MDarea.pop$TRACT + MDarea.pop$BLKGROUP
res1<-round(rbind(
  y1=BW3stagePPS(X=MDarea.pop$Hispanic, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y2=BW3stagePPS(X=MDarea.pop$Gender, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y3=BW3stagePPS(X=MDarea.pop$Age, pp=pp.trt1,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG)),digits = 3)
rownames(res1)<-c("Hispanic","Gender","Age")
res1

```

## b (three-stage(ppswr/srs/srs))
```{r}

pp.trt2 <- table(MDarea.pop$TRACT) / nrow(MDarea.pop) 
MDarea.pop$trtBG <- 10*MDarea.pop$TRACT + MDarea.pop$BLKGROUP
res2<-round(rbind(
  y1=BW3stagePPS(X=MDarea.pop$Hispanic, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y2=BW3stagePPS(X=MDarea.pop$Gender, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG),
y3=BW3stagePPS(X=MDarea.pop$Age, pp=pp.trt2,
psuID=MDarea.pop$TRACT, ssuID=MDarea.pop$trtBG)),digits = 3)

rownames(res2)<-c("Hispanic","Gender","Age")
res2
```


## c
For all three variables,when switching to a ppswr/srs/srs design from an srs/srs/srs design, $W^2$,$w_2^2$,and $W_3^2$ decreased slightly and $\delta_2^2$ increased slightly. However, the value of $B^2$ and $\delta_1^2$ became much smaller when PPS is used in the first stage instead of SRS. The $\delta_1^2$ is a measure of the homogeneity among the PSU totals. The decrease in $\delta_1^2$ indicates that the variation within PSUs is much larger than the variation among PSU totals under the ppswr/srs/srs design. This is most obvious in the variable "Hispanics" and "Gender".
