---
title: "SURVMETH Class 8 Lab Code"
author: "Stacey Frank"
date: "3/11/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(leaflet)
library(glue)
library(sf)
library(tidycensus)

options(tigris_class ="sf")
options(tigris_use_cache =T)

```

# Example 1: Create map of Washington Dc tracts, block groups and blocks

## Dc Tract Shapefile

```{r}

dc_trt <- tracts("District of Columbia", year=2019)
dc_trt_trans <- st_transform(dc_trt, 
                             "+proj=longlat +datum=WGS84")

```

```{r}
dc_trt[1:2,]
```
## DC Tract Map

```{r}
leaflet() %>%
  addTiles() %>%
  addPolygons(data=dc_trt_trans,
              popup = ~NAME)
```
## Blockgroup and block map shapefile

```{r}
dc_bg <- block_groups("District of Columbia", year=2019)
dc_bg_trans <- st_transform(dc_bg, "+proj=longlat +datum=WGS84")

dc_blk <- blocks("District of Columbia", year=2019)
dc_blk_trans <- st_transform(dc_blk, "+proj=longlat +datum=WGS84")
```

## DC blockgroup map layered with tracts

```{r}
leaflet() %>%
  addTiles() %>%
  addPolygons(data=dc_trt_trans,
              popup = ~NAME) %>%
  addPolygons(data=dc_bg_trans,
              popup = ~GEOID,
              color = "red",
              weight = 1,
              opacity= 1,
              smoothFactor = 1,
              fillOpacity = 0,
              fillColor= "#BDBDC3"
              ) 
```
## Now add Dc blocks layered with block groups and tracts

```{r}
leaflet() %>%
  addTiles() %>%
  addPolygons(data=dc_trt_trans,
              popup = ~NAME,
              fillOpacity = 0,
              fillColor= "snow") %>%
  addPolygons(data=dc_bg_trans,
              popup = ~GEOID,
              color = "red",
              weight = 2,
              opacity= 0.8,
              smoothFactor = 1,
              fillOpacity = 0,
              fillColor= "snow") %>%
  addPolygons(data=dc_blk_trans,
              popup = ~GEOID10,
              color = "green",
              weight = 0.7,
              opacity= 2,
              smoothFactor = 1,
              fillOpacity = 0,
              fillColor= "snow")
```
##Now are selecting only a few blockgroups (will need to do this for project)

```{r}
dc_bg_select<-dc_bg_trans [31:40,]

leaflet() %>%
  addTiles() %>%
  addPolygons(data=dc_trt_trans,
              popup = ~NAME) %>%
  addPolygons(data=dc_bg_select,
              color = "red",
              weight = 1,
              popup = ~GEOID)
```
# Example 2: To create a map of Washington DC tracts with ACS 2019 summary file data with

## Median income, population size, population proportion of blacks

```{r}
api_key <-"c3d3ea517c64b3c262d71c51135e449f6d98bf24" 
# Request at https://api.census.gov/data/key_signup.html 

census_api_key(key=api_key) 
# "install=T" stores your api_key
# If stored, checking api_key as follows:
#    readRenviron("~/.Renviron")
#    Sys.getenv("CENSUS_API_KEY")
```
```{r}
View(load_variables(2019, "acs5", cache = TRUE))
```

# Map with median income 
## B06011_001 is the median income

## First get income data from ACS

```{r}
DC_INCOME <- get_acs(geography = "tract", 
                     variables = "B06011_001", 
                     state = "District of Columbia",
                     geometry = TRUE) 

head(DC_INCOME)
```
## How I want to note the income differences on the map

```{r}
palette <- colorBin(palette = "Blues", 
                 domain = DC_INCOME$estimate, bins=10)

```

## More information about palette color options: http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3

## To create the income map

```{r}
DC_INCOME %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~GEOID,
              color = "black",
              weight = 1)%>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ palette(estimate)) %>%
  addLegend("bottomleft", 
            pal = palette, 
            values = ~ estimate,
            title = "Median income - Tract",
            labFormat = labelFormat(prefix = "$"),
            opacity = 1) 
```
## Map with population counts

### From the variable list, we know B01001_00 is the total population count and B01001B_001 is the count of Blacks

```{r}
DC_BLACK <- get_acs(geography = "tract", 
                    variables = "B01001B_001", 
                    state = "District of Columbia",
                    geometry = TRUE,
                    summary_var="B01001_001")  
```
### estimate is the counts of Blacks (B01001B_001) and symmary_est is the count of populaiton (B01001_001)

## Map of total population counts

```{r}

pal1 <- colorBin(palette = "Greys", 
                 domain = DC_BLACK$summary_est, bins=10)

DC_BLACK %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~GEOID,
              color = "black",
              weight = 1)%>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal1(summary_est)) %>%
  addLegend("bottomleft", 
            pal = pal1, 
            values = ~ summary_est,
            title = "Total population - Tract",
            opacity = 1)
```
## Map with population proportion of Blacks

```{r}
pal2 <- colorBin(palette = "BuGn", 
                 domain = DC_BLACK$estimate/DC_BLACK$summary_est*100, 
                 bins=10)

DC_BLACK %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~GEOID,
              color = "black",
              weight = 1)%>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal2(estimate/summary_est*100)) %>%
  addLegend("bottomleft", 
            pal = pal2, 
            values = ~ estimate/summary_est*100,
            title = "Proportion of Blacks - Tract",
            labFormat = labelFormat(suffix="%"),
            opacity = 1)
```

